---
title: "Simple document"
output: html_document
---

```{r}
library(tidyverse)
```

Let's import the data sets

```{r}
litters_df=
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", ""))
litters_df = 
janitor::clean_names(litters_df)

pups_df =
   read_csv("./data/FAS_pups.csv", na = c("NA", "."), skip = 3)

pups_df = 
    janitor::clean_names(pups_df)
```

select a range

```{r}
select(litters_df, group)
```

```{r}
select(litters_df, -group)
select(litters_df, -group:gd_of_birth)
```

select by prefix

```{r}
select(litters_df, group, starts_with("gd"))
```

rearrange by selecting

```{r}
select(litters_df, starts_with("gd"), group)
select(litters_df, litter_number, everything())

relocate(litters_df, litter_number)
```

rename by selecting

```{r}
select(litters_df, GROUP = group, everything())
rename(litters_df, GROUP = group)
```

what if I only select one variable

```{r}
select(litters_df, group)

```

LA  code:

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter`

use `filter()` to filter out rows

```{r}
filter(litters_df, gd_of_birth == 20, pups_born_alive == 6)
```

look for more than two pups alive

```{r}
filter(litters_df, pups_born_alive > 5)
filter(litters_df, pups_born_alive >= 5)
filter(litters_df, pups_born_alive <= 5)
```

look for non matches

```{r}
filter(litters_df, pups_born_alive != 6)
```

match to a set

```{r}
filter(litters_df, group %in% c("Con7", "Con8"))
```

filter out missing rows with `drop_na()`

```{r}
drop_na(litters_df)
drop_na(litters_df, gd0_weight)
```

filter to include only pups with sex1
filter to include only pups with PD walk less than 11 and sex2

```{r}
filter(pups_df, sex == 1)
filter(pups_df, sex == 2, pd_walk < 11 )
```

## `mutate`

use `mutate()` to create or modify variables

```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```


